Portfolios of Subgraph Isomorphism Algorithms

### 2016-04-12 paper/paper.tex
Christine's changes

The subgraph isomorphism problem is to find an adjacency-preserving injective mapping from vertices
of a small \emph{pattern} graph to vertices of a large \emph{target} graph. This \NP-complete
problem has many important practical applications, for example in computer
vision~\cite{cviu11,pr15}, biochemistry~\cite{Giugno:2013}, and model checking~\cite{Sevegnani:2015}. There
exist various exact algorithms, which have been compared on a large suite of problem instances by
McCreesh and Prosser~\cite{McCreesh:2015}. These experiments indicated that the single best
algorithm depends on the CPU time limit considered: for very small CPU time
limits, \VFtwo{}~\cite{Cordella:2004} is the best choice, whereas the \Glasgow algorithm~\cite{McCreesh:2015} has
better success rates for larger time limits.  They also showed that on an instance by instance
basis, other algorithms are often better.

@@

Here, we consider the case where exactly one algorithm is selected for solving the problem. One of
the most prominent and successful systems that employs this approach is
SATzilla~\cite{xu_satzilla_2008}, which defined the state of the art in SAT solving for a number of
years. Other application areas include constraint solving~\cite{omahony_using_2008}, the travelling
salesperson problem~\cite{kotthoff_improving_2015}, and AI planning~\cite{seipp_learning_2012}.
The interested reader is referred to a recent survey~\cite{kotthoff_algorithm_2014} for additional
information on algorithm selection.

\paragraph{Overview of the paper.}
We formally define the subgraph isomorphism problem in Section~\ref{sec:defs}.
In Section~\ref{sec:algs}, we describe the main existing algorithms for solving
this problem, and we also introduce two new algorithms which are derived from
Solnon's \LAD algorithm \cite{Solnon:2010}.

In Section~\ref{sec:exps}, we experimentally compare eight state-of-the-art
algorithms. We introduce a large benchmark set composed of 5725 instances
grouped into twelve classes. Ten of these classes were considered in the
experimental study reported by McCreesh and Prosser~\cite{McCreesh:2015}; two
are new. We evaluate the algorithms on this benchmark set, and show that they
have very complementary performance. In particular, we show that depending on
the CPU time limit, different algorithms achieve the best performance on the
entire benchmark set.

In Section~\ref{sec:algsel}, we discuss the features that are used to describe
instances, and we describe our algorithm selection approach. It combines a
presolving step, which allows us to easy instances very quickly, with an
algorithm selection step that uses \LLAMA~\cite{kotthoff_llama_2013}.
In Section~\ref{sec:algsel-exps}, we experimentally evaluate our selection
approach and show that it is able to close more than \SI{60}{\percent} of the
gap between the single best and the virtual best solver. We conclude and give
directions for future work in Section~\ref{sec:concs}.

@@

Given a pattern graph $G_p=(N_p,E_p)$ and a target graph $G_t=(N_t,E_t)$, the \emph{subgraph
isomorphism problem} consists of deciding whether $G_p$ is isomorphic to some subgraph of $G_t$.
More formally, the goal is to find an injective matching $f: N_p\rightarrow N_t$, that associates a
different target node to each pattern node, and preserves pattern edges, i.e.\ $\forall (u,u')
\in E_p, (f(u),f(u')) \in E_t$.
Note that the subgraph is not necessarily induced, so that two pattern nodes not linked by
an edge may be mapped to two target nodes which are linked by an edge.

We define $n_p = \# N_p$, $n_t = \# N_t$,  $e_p=\# E_p$, $e_t=\# E_t$, and $d_p$
and $d_t$ to be the maximum degrees of the graphs $G_p$ and $G_t$.

@@

We selected the following algorithms from the literature, based on their
performance.
\begin{itemize}
\item \VFtwo{}~\cite{Cordella:2004} performs weak filtering that is especially fast on
    trivially satisfiable instances;
\item \LAD{}~\cite{Solnon:2010} combines two strong but expensive filtering techniques
    (All-Different Generalized Arc Consistency and Locally All-Different);
\item \Glasgow{}~\cite{McCreesh:2015} does expensive preprocessing based on path length
    invariant properties to generate additional constraints, followed by weaker filtering
    (forward-checking, and a heuristic All-Different propagator which can miss deletions) and
    conflict-directed backjumping during search.
\end{itemize}

@@

\begin{enumerate}
    \item If \IncompleteLAD presolving removes at least
        \SI{28.014755}{\percent} of the pairs, and at least
        \SI{94.117645}{\percent} of the values from at least one domain, then
        pick \PathLAD.
    \item If the target has at least $610$ vertices, and if the maximum
        distance between any two pattern vertices is at most $8$, and if the
        pattern is not regular, and if the time taken to compute the
        distance-based features on the target graph is no more than
        \SI{1277}{\ms}, then pick \PathLAD.
    \item If \IncompleteLAD filtering removed at least \SI{5.904644}{\percent} of
        possible pairs, and if less than \SI{84.6565}{\percent} of the pattern
        vertices are within distance 2 of each other, then pick \PathLAD.
    \item Otherwise, pick \GlasgowTwo.
\end{enumerate}

>>>>

The subgraph isomorphism problem is to find an adjacency-preserving injective mapping from vertices
of a small \emph{pattern} graph to vertices of a large \emph{target} graph. This \NP-complete
problem has many important practical applications, for example in computer
vision~\cite{cviu11,pr15}, biochemistry~\cite{Giugno:2013}, and model checking~\cite{Sevegnani:2015}. There
exist various exact algorithms, which have been compared on a large suite of instances by
McCreesh and Prosser~\cite{McCreesh:2015}. These experiments indicated that the single best
algorithm depends on the CPU time limit considered: for very small time
limits, \VFtwo{}~\cite{Cordella:2004} is the best choice, whereas the \Glasgow algorithm~\cite{McCreesh:2015} has
better success rates for larger time limits.  They also showed that on an instance by instance
basis, other algorithms are often better.

@@

% MODIFICATIONS
In our subgraph isomorphism context, algorithm performance is highly constrained by memory bandwidth (as pointed out by Sabharwal and Samulowitz \cite{Sabharwal2014} for SAT solvers).
Therefore, we cannot simply run different algorithms in parallel, and we consider the case where exactly one algorithm is selected for solving the problem. 
%
One of
the most prominent and successful systems that employs this approach is
SATzilla~\cite{xu_satzilla_2008}, which defined the state of the art in SAT solving for a number of
years. Other application areas include constraint solving~\cite{omahony_using_2008}, the travelling
salesperson problem~\cite{kotthoff_improving_2015}, and AI planning~\cite{seipp_learning_2012}.
The interested reader is referred to a recent survey~\cite{kotthoff_algorithm_2014} for additional
information on algorithm selection.

\paragraph{Overview of the paper.}
We formally define the subgraph isomorphism problem in Section~\ref{sec:defs}.
In Section~\ref{sec:algs}, we describe the main existing algorithms for solving
this problem, and we also introduce two new algorithms which are derived from
Solnon's \LAD algorithm \cite{Solnon:2010}.
In Section~\ref{sec:exps}, we experimentally compare eight state-of-the-art
algorithms. We introduce a large benchmark set composed of 5725 instances
grouped into twelve classes. Ten of these classes were considered in the
experimental study reported by McCreesh and Prosser~\cite{McCreesh:2015}; two
are new. We evaluate the algorithms on this benchmark set, and show that they
have very complementary performance. In particular, we show that depending on
the CPU time limit, different algorithms achieve the best performance on the
entire benchmark set.
In Section~\ref{sec:algsel}, we discuss the features that are used to describe
instances, and we describe our algorithm selection approach. It combines a
presolving step, which allows us to easy instances very quickly, with an
algorithm selection step that uses \LLAMA~\cite{kotthoff_llama_2013}.
In Section~\ref{sec:algsel-exps}, we experimentally evaluate our selection
approach and show that it is able to close more than \SI{60}{\percent} of the
gap between the single best and the virtual best solver. We conclude and give
directions for future work in Section~\ref{sec:concs}.

@@

Given a pattern graph $G_p=(N_p,E_p)$ and a target graph $G_t=(N_t,E_t)$, the \emph{subgraph
isomorphism problem} consists of deciding whether $G_p$ is isomorphic to some subgraph of $G_t$.
More formally, the goal is to find an injective matching $f: N_p\rightarrow N_t$, that associates a
different target node to each pattern node, and preserves pattern edges, i.e.\ $\forall (u,u')
\in E_p, (f(u),f(u')) \in E_t$.
Note that the subgraph is not necessarily induced, so that two pattern nodes not linked by
an edge may be mapped to two target nodes which are linked by an edge.
We define $n_p = \# N_p$, $n_t = \# N_t$,  $e_p=\# E_p$, $e_t=\# E_t$, and $d_p$
and $d_t$ to be the maximum degrees of the graphs $G_p$ and $G_t$.

@@

We selected the following algorithms from the literature, based on their
performance.
\begin{itemize}
\item \VFtwo{}~\cite{Cordella:2004} performs weak filtering that is especially fast on
    trivially satisfiable instances;
\item \LAD{}~\cite{Solnon:2010} combines two strong but expensive filtering techniques
    (All-Different Generalized Arc Consistency and Locally All-Different);
\item \Glasgow{}~\cite{McCreesh:2015} does expensive preprocessing based on path length
    invariant properties to generate additional constraints, followed by weaker filtering
    (forward-checking, and a heuristic All-Different propagator which can miss deletions) and
    conflict-directed backjumping during search.
\end{itemize}
% MODIFICATIONS
We have not considered the algorithm introduced in \cite{constraints10} because it is outperformed by \LAD{}.
Also, we have not considered MIP nor SAT solvers because they are not competitive with the selected algorithms \cite{UpcomingIJCAIPaper}.
%

@@

% MODIFICATIONS: removed decimals (just keep to digits after the point)
\begin{enumerate}
    \item If \IncompleteLAD presolving removes at least
        \SI{28.01}{\percent} of the pairs, and at least
        \SI{94.12}{\percent} of the values from at least one domain, then
        pick \PathLAD.
    \item If the target has at least $610$ vertices, and if the maximum
        distance between any two pattern vertices is at most $8$, and if the
        pattern is not regular, and if the time taken to compute the
        distance-based features on the target graph is no more than
        \SI{1277}{\ms}, then pick \PathLAD.
    \item If \IncompleteLAD filtering removed at least \SI{5.90}{\percent} of
        possible pairs, and if less than \SI{84.66}{\percent} of the pattern
        vertices are within distance 2 of each other, then pick \PathLAD.
    \item Otherwise, pick \GlasgowTwo.
\end{enumerate}
